Thermoacoustic heat pumps

### 2023-04-26 paper.tex
added automotive

\documentclass{article}
\usepackage{graphicx} % Required for inserting images
\usepackage{enumerate}% http://ctan.org/pkg/enumerate
\usepackage[superscript,biblabel]{cite}

@@

\subsection{Domestic applications}
Almost 80\% of domestic heating in Europe is still being done with fossil fuels according to the EU\cite{EU}. There are two startups in Europe seeking to improve these numbers by making heat pumps a more accessible and environment-friendly option.
\subsection{Blue Heart\cite{blueheart}}
Blue Heart is a Dutch spinoff of the Organisation for Applied Scientific Research (TNO). A Blue Heart device would replace the cold circuit of a heat pump. High energy density of TAHP reduces the space needed for heat pump installations. They developed a traveling wave TAHP with helium as the working fluid. The acoustic power is generated by two pistons opposite to each other moving at the same frequency. The vibrations get canceled making the TAHP more silent than their conventional compression-based counterparts. Blue Heart claims that its device works efficiently with every temperature input and output and has a larger operation envelope than compression-based heat pumps. They also claim to be compatible with every type of heat source and to be suitable for new and existing houses.
\newpara{}
A Blue Heart device would operate with source temperatures between -20 degrees to 50 degrees Celsius and with sink temperatures between 7 degrees and 80 degrees Celsius. It would also have a capacity of 6 kW. The company plans to make the device commercially available in 2024.
\subsection{Equium\cite{equium}}
Equium is a French startup that offers to replace the heat pump entirely in contrast to Blue Heart. The company has not revealed the inner workings of their device but it is expected to also be a traveling wave TAHP working with helium. The company promises an impressive maintenance-free lifetime of 30 years. These devices are also planned to be commercially available in 2024
\section{Future challenges and limitations}

>>>>

\documentclass[a4paper]{article}
\usepackage{graphicx} % Required for inserting images
\usepackage{enumerate}% http://ctan.org/pkg/enumerate
\usepackage[superscript,biblabel]{cite}

@@

\subsection{Domestic applications}
Almost 80\% of domestic heating in Europe is still being done with fossil fuels according to the EU\cite{EU}. There are two startups in Europe seeking to improve these numbers by making heat pumps a more accessible and environment-friendly option.
\subsubsection{Blue Heart\cite{blueheart}}
Blue Heart is a Dutch spinoff of the Organisation for Applied Scientific Research (TNO). A Blue Heart device would replace the cold circuit of a heat pump. The high energy density of TAHP reduces the space needed for heat pump installations. They developed a traveling wave TAHP with helium as the working fluid. The acoustic power is generated by two pistons opposite to each other moving at the same frequency. The vibrations get canceled making the TAHP more silent than their conventional compression-based counterparts. Blue Heart claims that its device works efficiently with every temperature input and output and has a larger operation envelope than compression-based heat pumps. They also claim to be compatible with every type of heat source and to be suitable for new and existing houses.
\newpara{}
A Blue Heart device would operate with source temperatures between -20 degrees to 50 degrees Celsius and with sink temperatures between 7 degrees and 80 degrees Celsius. It would also have a capacity of 6 kW. The company plans to make the device commercially available in 2024.
\subsubsection{Equium\cite{equium}}
Equium is a French startup that offers to replace the heat pump entirely in contrast to Blue Heart. The company has not revealed the inner workings of their device but it is expected to also be a traveling wave TAHP working with helium. The company promises an impressive maintenance-free lifetime of 30 years. These devices are also planned to be commercially available in 2024
\subsection{Automotive applications}
Current refrigeration and cooling systems in automotive applications, based on compression, draw a lot of power from the engine and degrades the overall efficiency of the vehicle. Furthermore, the refrigerant often leaks from the system and is made of ozone-depleting material. An environment-friendly and efficient alternative to cooling systems in vehicles is thus crucial.
\newpara{}
Zoontjens et al.\ did a feasibility study on TAHP in automotive applications\cite{zoontjens2005feasibility}. They specifically studied the possibility of a thermally driven TAHP driven by recovered waste heat from waste exhaust gasses. They also compared this alternative to other prominent alternatives such as active magnetic resonator systems, solid adsorption cooling systems, vapor absorption refrigeration systems, etc.  What makes TAHP appealing, in this case, is the possibility to be driven thermally because combustion engines have low thermal efficiencies and thus release a significant amount of heat to the environment. A diagram of the system they devised is shown in Figure~\ref{automotiveDiagram} (ATAR stands for automotive thermoacoustic refrigeration).
\begin{figure}[h]
  \centering
  \includegraphics[width=0.5\textwidth]{images/automotive.png}
  \caption{Diagram of automotive thermoacoustic refrigerator\cite{zoontjens2005feasibility}}\label{automotiveDiagram}
\end{figure}
The implementation chosen for this study was a standard quarter-wavelength standing wave TAHP as discussed above. The two major concerns are cooling power and available heating power. The cost of developing a production-ready system of this kind is a big challenge to overcome. Especially with the ever-decreasing demand for combustion engine vehicles. The researchers found that these systems would be feasible but that further optimization and research will be necessary.
\section{Future challenges and limitations}
